

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{tcolorbox}
\tcbuselibrary{theorems}
\usepackage{tikz-cd}
\usepackage{xparse}
\usepackage{adjustbox}
\usepackage{braket}



\usepackage{tabularx}

%\newcommand{\cat}[1]{\mathcal{#1}}



\title{The Wightman Axioms for Hopf-Frobenius Modules}

\author{Matt Alexander}
\usetheme{queer}


%\newcommand{\newthought}[1]{\medskip\noindent\fbox{{#1}}}


\input{definitions.tex}



\input{usepackage}



\begin{document}


\begin{frame}
\titlepage
\end{frame}



%===================
% SLIDE 1
%===================







%===================
% SLIDE 2
%===================



\begin{frame}

\frametitle{Definition of Frobenius Algebras}

%\framesubtitle{Part 3 --- Equivalences}

A Frobenius algebras is a unital associative $k$-algebra with a pairing $\gen{\cdot, \cdot}: F \otimes F \to k$, such that

\begin{itemize}
\item \newthought{Non-degenerate: } If $\gen{x,y} = 0$ for all $y$, then $x = 0$. Similarly, $\gen{x,y} = 0$ for all $x$ implies $y=0$.

\item \fbox{Associativity: } $\gen{xy,z} = \gen{x,yz}$, or in the complex case, $\gen{xy,z} = \gen{x, y^* z}$.
\end{itemize}


\newthought{Equivalently: } A linear functional $\varepsilon: F \to k$, that satisfies the non-degeneracy condition
\begin{equation*}
\gen{x,y} = \varepsilon(xy) \quad \gen{x,y} = \varepsilon(x^* y)
\end{equation*}


\end{frame}


%===================
% SLIDE 1
%===================



%===================
% SLIDE 2
%===================



\begin{frame}
\frametitle{Frobenius Algebras}

\framesubtitle{Examples}



\begin{tabularx}{\textwidth}{lll>{\raggedright\arraybackslash}X}
\toprule
Space & $\mu$ & $\gen{\cdot, \cdot}$ 
\\
\midrule
\\
$C_c^{\infty}(X)$ & Pointwise & $\gen{f,g} = \int f^* g$ \\
Schwartz functions $\cat{S}$ & Pointwise & $\gen{f,g} = \int f^* g$
\\
\\
$\mathbb{R}^n$ w/ basis $\{ e_i \}$ & $e_i e_j = \delta_{ij} e_i$ & $\gen{e_i,e_j} = \delta_{ij}$ \\
Hilbert space w/ basis $\{ e_i \}$ & $e_i e_j = \delta_{ij} e_i$ & $\gen{e_i,e_j} = \gen{e_i, e_j}_H$  \\
\\
$kG$ (finite group) & Group mult. & $\gen{g,h} = \delta_{gh, 1}$ 
\\
\bottomrule
\end{tabularx}

\medskip


\begin{equation*}
\cat{S} = \{ f \in C^{\infty}(\mathbb{R}^n, \mathbb{C}) \mid \sup_{x} \abs{x^\alpha D^\beta f(x)} < \infty, \forall \alpha, \beta \}
\end{equation*}



\end{frame}

%===================
% SLIDE 3
%===================



\begin{frame}
\frametitle{Hopf Algebras}

\framesubtitle{Definition}

A Hopf algebra is a unital associative $k$-algebra, $H$, with maps 

\medskip

\begin{itemize}
\item \fbox{Comultiplication: } $\Delta: H \to H \otimes H$

\item \fbox{Counit: } $\varepsilon: H \to k$

\item \fbox{Antipode: } An involution $S: H \to H$
\end{itemize}
which satisfy appropriate compatibility conditions.

\medskip

\adjustbox{center}{
\begin{tikzpicture}
  \node {$H$}
    child {node {$\mu$}
    		child {node {$H$}}
    		child {node {$H$}}
    	};
\end{tikzpicture}
\quad
\raisebox{6pt}{
\begin{tikzpicture}
  \node {$H$}
    child {node {$\eta$}
    		child {[fill] circle (2pt)}
    	};
\end{tikzpicture}}
\quad
\begin{tikzpicture}
  \node {$H$} [grow'=up]
    child {node {$\Delta$}
    		child {node {$H$}}
    		child {node {$H$}}
    	};
\end{tikzpicture}
\quad
\begin{tikzpicture}[grow'=up]
  \node {$H$}
    child {node {$\varepsilon$}
    		child {[fill] circle (2pt)}
    	};
\end{tikzpicture}
\quad
\begin{tikzpicture}
  \node {$H$}
    child {node {$S$}
    		child {node {$H$}}
    	};
\end{tikzpicture}
}

%
%\begin{tikzpicture}
%  \node[fill, circle] {}
%    child {[fill] circle (2pt)}
%    child {[fill] circle (2pt)} {child {[fill] circle (2pt)}}
%    child {[fill] circle (2pt)};
%\end{tikzpicture}

\end{frame}










\begin{frame}


\frametitle{Hopf Algebras}
%

\framesubtitle{Examples}

\begin{tabularx}{\textwidth}{lll>{\raggedright\arraybackslash}X}
\toprule
Space & $\Delta$ & $\varepsilon$
\\
\midrule
\\
$kG$  & $\Delta g = g \otimes g$ & $\varepsilon(g) = 1$ \\
\\
Tensor alg. $T(X)$ & $\Delta x = x \otimes 1 + 1 \otimes x$ & $\varepsilon(x) = 0$ \\
Symmetric alg. $S(X)$ & $\Delta x = x \otimes 1 + 1 \otimes x$ & $\varepsilon(x) = 0$ \\
Univ. Env. Alg. $U(\mathfrak{g})$ & $\Delta x = x \otimes 1 + 1 \otimes x$ & $\varepsilon(x) = 0$
\\
\\
$\Lambda(X)$ & $\Delta x = x \otimes 1 - 1 \otimes x$ & $\varepsilon(x) = 0$
\\
\bottomrule
\end{tabularx}

\medskip 

\newthought{Sweedler Notation: } Denote the comultiplication by
\begin{equation*}
\Delta x = \sum x_{(1)} \otimes x_{(2)}.
\end{equation*}

\end{frame}

%===================
% SLIDE 4
%===================


\begin{frame}

\frametitle{Hopf-Frobenius Modules}
\framesubtitle{Definition}

A left Hopf-Frobenius module is a pair $(H,F)$ where $H$ is a Hopf algebra, $F$ is a Frobenius algebra, such that the following conditions are satisfied:

\bigskip

\begin{itemize}

\item \fbox{Module: } $F$ is a left $H$-module.

\medskip

\item \fbox{Weak Cartan: }  $\sum \gen{g_{(1)} \cdot x, g_{(2)} \cdot y} = \varepsilon(g) \gen{x,y}$.

\medskip

\item \fbox{Antipode Condition: }  $\gen{S(g) \cdot x, y} = \gen{x, g \cdot y}$.

\end{itemize}





\end{frame}


\begin{frame}

\frametitle{Hopf-Frobenius Modules}
\framesubtitle{Example 1 --- Derivatives of Functions}


Take our Hopf algebra to be $k[\partial]$ and the Frobenius algebra to be $\cat{S}$, the space of Schwartz functions. This is an HF-module with

\bigskip

\begin{itemize}

\item \fbox{Action: } $\partial \cdot f = f'$, the derivative of $f$


\medskip

\item \fbox{Weak Cartan / Antipode: } Both of these conditions reduce to

\begin{equation*}
\int (\partial \cdot f) g  = - \int f (\partial \cdot g)
\end{equation*}
The \textbf{weak derivative} condition.
% $\gen{\partial f, g} + \gen{f, \partial g} = \varepsilon(\partial) \gen{f,g} = 0$. So
%\begin{equation}
%\int (\partial f) g + \int f (\partial g) = 0
%\end{equation}
%
%\item \fbox{Antipode: } $S(\partial) = - \partial$
\end{itemize}


\end{frame}



\begin{frame}

\frametitle{Hopf-Frobenius Modules}
\framesubtitle{Example 2 --- Operators on a Hilbert Space}

Let $F$ be a separable Hilbert space, and let $\{ A_i \}$ be a family of compact mutually-commuting \textit{skew-Hermitian} operators on $F$. Then we can find an orthonormal basis of simultaneous eigenvectors, $\{ e_k \}$. 

Take $(F, \{ e_k \})$ as our Frobenius algebra, and let $H = U(\{ A_i \} )$ be the universal enveloping algebra of the abelian Lie algebra generated by the $\{ A_i \}$. This is a HF-module where


\begin{itemize}

\item \fbox{Action: } $A_i \cdot e_k = A_i e_k = \alpha_{i}^k e_k$, the natural action of the Hermitian operators on the eigenvector basis.


\medskip

\item \fbox{Weak Cartan / Antipode: } Both of these conditions reduce to

\begin{equation*}
\gen{H x, y} = - \gen{x, Hy}
\end{equation*}
The \textbf{skew-Hermitian} condition.
\end{itemize}



\end{frame}





\begin{frame}

\frametitle{Hopf-Frobenius Modules}
\framesubtitle{Example 3 --- Lie Groups}


Consider the Lie group $U(n)$, and let $H$ be its group algebra over $\mathbb{C}$. Take $\mathbb{C}^n$ with basis $\{ e_i \}$ as our Frobenius algebra.

\bigskip

\begin{itemize}

\item \fbox{Action: } Left multiplication.


\medskip

\item \fbox{Weak Cartan / Antipode: } Both of these conditions reduce to

\begin{equation*}
\gen{g \cdot x, g \cdot y} = \gen{x,y}
\end{equation*}
The unitarity condition.

\end{itemize}




\end{frame}



%===================
% SLIDE 2
%===================



\begin{frame}


\frametitle{Hopf-Frobenius Modules}
%

\begin{tabularx}{\textwidth}{ll|ll>{\raggedright\arraybackslash}X}
\toprule
Frobenius & & Hopf
\\
\midrule
& & \\
States & \textbf{Hil} & Operators & $B(\textbf{Hil})$ \\
Functions & $\cat{S}$ & Derivatives & $\mathbb{C}[\partial_1, \ldots, \partial_n]$ \\
Space & $\mathbb{R}^{1,3}$ & Transformations & $\Lambda$
\\
&& \\
\bottomrule
\end{tabularx}



\end{frame}



%===================
% SLIDE 2
%===================




\begin{frame}
\frametitle{Wightman Axioms}
\framesubtitle{Relativistic Quantum Mechanics}


\begin{enumerate}
\item \fbox{Hilbert Space: } There is a separable Hilbert space of states $X$.

\item \fbox{Evolution: } There is a unitary representation of $SL(2, \mathbb{C})$ on $X$, which we denote $U(L,a)$

\item \fbox{Spectral Condition: } We can write $U(1,a) = \exp(i P^{\mu} a_\mu)$ and the eigenvalues of $P$ satisfy $p_0^2 - \sum_{i=1}^3 p_i^2 > 0$ and $p^0 > 0$.

\item \fbox{Vacuum: } There is a unique vector $\ket{0} \in X$ invariant under $U(L,a)$.
\end{enumerate}


\end{frame}





%===================
% SLIDE 2
%===================


\begin{frame}

\frametitle{Wightman Axioms}
\framesubtitle{Field Conditions}

\begin{enumerate}
\item \fbox{Dense Subset: } There is a dense subset $D \subseteq X$ which contains $\ket{0}$ and is mapped to itself by $U(L,a)$.

\item \fbox{Fields: } Quantum fields are operator-valued distributions on a Schwartz space, $\phi_i: \cat{S} \to B(D)$.

\item \fbox{Transformations: } There is an action of $SL(2)$ on the fields, $S(L)$ such that $U(L,a) \phi(x) U(L,a)^{-1} = S(L) \phi(L^{-1}(x - a))$

\item \fbox{Cyclic Vacuum: } The image of polynomials in $\phi_i(f)$ on $\ket{0}$ is dense in $X$.
\end{enumerate}


\end{frame}


%===================
% SLIDE 2
%===================


\begin{frame}

\frametitle{Wightman Axioms}
\framesubtitle{Microcausality Condition}

\begin{enumerate}

\item If $f,g \in \cat{S}$ have spacelike supports ($f(x) g(y) = 0$ for all $x,y$ with $d(x,y) \geq 0$) then $[\phi_i (f), \phi_j (g)]_{\pm}(v) = 0$ for all $v \in D$.

%[spacelike is negative distance]

\end{enumerate}

\end{frame}






%===================
% SLIDE 2
%===================




\begin{frame}
\frametitle{Hopf-Frobenius Wightman Axioms}
%\framesubtitle{Relativistic Quantum Mechanics}

A (weak) Hopf-Frobenius-Wightman QFT is a collection of Hopf algebras ($B(F_{\text{state}}), k SL(2)$)  and Frobenius algebras ($F_{\text{state}}, \cat{S}, \mathbb{R}^{1,3}$)  such that

\begin{enumerate}


\item \fbox{States and Evolution: } $(k SL(2),F_{\text{state}})$ forms a Hopf-Frobenius module.

\item \fbox{Vacuum: } There is a unit $1 \in F_{\text{state}}$, which is the unique element preserved by $k SL(2)$.

\item \fbox{Fields: } Quantum fields are linear maps $\cat{S} \to B(F_{\text{state}})$ and form a Frobenius algebra $F_{\text{field}}$.

\item \fbox{Field Transformations: } $(k SL(2),F_{\text{field}})$ forms a Hopf-Frobenius module.





\end{enumerate}


\end{frame}




\begin{frame}
\frametitle{Hopf-Frobenius Wightman Axioms}
\framesubtitle{Other Conditions}

\begin{enumerate}

\item \fbox{Dense Subset: } We can view $F_{\text{state}}$ as all of $D$.


\item \fbox{Cyclic Vacuum: } The image of polynomials in $\phi_i(f)$ on $\ket{0}$ is dense in $X$.


\item \fbox{Spectral Condition: } We can write $U(1,a) = \exp(i P^{\mu} a_\mu)$ and the eigenvalues of $P$ satisfy $p_0^2 - \sum_{i=1}^3 p_i^2 > 0$ and $p^0 > 0$.


\item \fbox{Causality: }  If $f,g \in \cat{S}$ have spacelike supports ($f(x) g(y) = 0$ for all $x,y$ with $d(x,y) \geq 0$) then $[\phi_i (f), \phi_j (g)]_{\pm}(v) = 0$ for all $v \in D$.


\end{enumerate}
\end{frame}

































\end{document}
%
%
%
%
%
%
